---
title: "Analysis_Code"
output: html_document
date: "2024-11-20"
---

To Do:

Extract data from API 
Categorize the neighborhoods by the criteria

Julia: grocery store distance
Amia: vehicles
Rosie: income

### Extract Data from API:  

HFAI  

```{r}
library(jsonlite)
library(tidyverse)

hfai_url <- "https://services1.arcgis.com/mVFRs7NF4iFitgbY/arcgis/rest/services/Hfai/FeatureServer/0/query?where=1%3D1&outFields=*&outSR=4326&f=json"

jsonData <- read_json(hfai_url, simplifyVector = TRUE)

hfai <- jsonData$features$attributes %>% 
  select(OBJECTID, CSA2010, hfai15, Shape__Area, Shape__Length)
head(hfai)


# LOW TIER IS 0-9.5
# MEDIUM TIER IS 9.6-19
# HIGH TIER IS 19.1 - 28.5
```
```{r}
hfai_for_joining <- hfai %>% 
  select(CSA2010, hfai15) %>% 
  mutate(hfai_category = case_when(
    hfai15 < 9.6 ~ "Low",
    hfai15 > 9.5 & hfai15 <= 19 ~ "Medium",
    hfai15 > 19 ~ "High"
  ), hfai_boolean = ifelse(hfai15 < 9.6, 1, 0))
hfai_for_joining
```


Median Income
Notes - use threshold of federal poverty level for a family of 4 ($31,200 in 2024). classify neighbourhoods as 1 where median income is 185% of this or less (in this case $57,720)

```{r}
library(jsonlite)
library(tidyverse)

income_url <- "https://services1.arcgis.com/mVFRs7NF4iFitgbY/arcgis/rest/services/Mhhi/FeatureServer/0/query?outFields=*&where=1%3D1&f=geojson"

jsonData <- read_json(income_url, simplifyVector = TRUE)

med_income <- jsonData$features$properties %>% 
  select(OBJECTID, CSA2010, mhhi22, Shape__Area, Shape__Length)

income_table_for_joining <- med_income %>%
  mutate(income_boolean = ifelse(mhhi22 < 57720, 1, 0)) %>%
  select(CSA2010, income_boolean, median_income = mhhi22)
```

Percent of Vehicles Available

```{r}
library(jsonlite)
library(tidyverse)

vehicles_url <- "https://services1.arcgis.com/mVFRs7NF4iFitgbY/arcgis/rest/services/Novhcl/FeatureServer/0/query?where=1%3D1&outFields=*&outSR=4326&f=json"

jsonData <- read_json(vehicles_url, simplifyVector = TRUE)

vehicles <- jsonData$features$attributes %>%
select(OBJECTID, CSA2010, novhcl22, Shape__Area, Shape__Length)

vehicle_table_for_joining <- vehicles %>%
  mutate(vehicle_boolean = ifelse(novhcl22 > 30, 1, 0)) %>%
  select(CSA2010, vehicle_boolean, percnt_no_vehicle = novhcl22)
head(vehicle_table_for_joining)

```

Grocery Store Location

```{r}
library(jsonlite)
library(tidyverse)

# install.packages("pak")
library(pak)
library(purrr)
#pak::pkg_install("elipousson/mapbaltimore", dependencies = TRUE)
library(mapbaltimore)
grocery_store_url <- "https://services1.arcgis.com/UWYHeuuJISiGmgXx/arcgis/rest/services/Grocery_Store/FeatureServer/0/query?where=1%3D1&outFields=*&outSR=4326&f=json"

jsonData <- read_json(grocery_store_url, simplifyVector = TRUE)

grocery_stores <- jsonData$features$attributes

# remove store in row 46 because it is outside of the zone and doesn't register as an address

add <- grocery_stores$address[1:45]

csas_with_store <- map(.x = add, .f = ~ get_area(type="csa", location = .x)$name)

# assign manually stores which did not get assigned a CSA

csas_with_store[[19]] <- "South Baltimore"
csas_with_store[[38]] <- "South Baltimore" 

csas_with_store <- unlist(csas_with_store)

grocery_stores_2 <- cbind(grocery_stores[1:45,], csas_with_store) %>% 
  select(OBJECTID, latitude, longitude, address, storename, csas_with_store) %>% 
  add_count(csas_with_store, name = "store_number") %>% 
  group_by(csas_with_store) %>% 
  mutate(store_number = 1:n()) %>% 
  pivot_wider(id_cols = csas_with_store, names_from = store_number, values_from = c(address, latitude, longitude, storename)) %>% 
  mutate(has_store = 1)
```
```{r}
# create vector with CSA names

csas <- data.frame(csa = vehicles$CSA2010)

grocery_store_join <- left_join(csas, grocery_stores_2, by = join_by(csa == csas_with_store)) 
grocery_store_join <- grocery_store_join %>%
  mutate(has_store = ifelse(is.na(grocery_store_join$has_store),0,1)) %>% 
  relocate(has_store, .after = csa) %>% 
  rename(CSA2010 = csa)
head(grocery_store_join)
```


#```{r}
contains_store <- function(n) {
  if(n %in% csas_with_store) {
    1
  } else {
    0
  }
}

grocery_stores_csa <- data.frame(cbind(csas, map_dbl(.x = csas, .f = ~contains_store(.x))))
colnames(grocery_stores_csa) <- c("CSA2010", "groc_store")
#grocery_stores_csa

# grocery store 1 or 0, grocery store #1 address, grocery sore #1 name, repeat for #2-#4

#data.frame(cbind(mapbaltimore::csas$name, vehicles$CSA2010))
#vehicles$CSA2010 %in% mapbaltimore::csas$name
#length(mapbaltimore::csas$name)
#length(vehicles$CSA2010)
mapbaltimore::csas$name %in% vehicles$CSA2010
#mapbaltimore::csas$name[51]

### Which areas have over 30% of households without a vehicle available?


### Joining income, vehicle, HFAI grocery store, and location data
```{r}
master_neighbourhood_data <- income_table_for_joining %>%
  left_join(vehicle_table_for_joining, by = 'CSA2010') %>%
  left_join(grocery_store_join, by = 'CSA2010') %>%
  left_join(hfai_for_joining, by = 'CSA2010') %>%
  mutate(
    number_criteria_met = income_boolean + vehicle_boolean + has_store + hfai_boolean,
    priority_level = case_when(
      number_criteria_met == 4 ~ 'High',
      number_criteria_met > 2 ~ 'Medium',
      number_criteria_met < 3 ~ 'Low'
    )
  )
master_neighbourhood_data
```

```{r}
write_csv(master_neighbourhood_data, "master_neighbourhood_data.csv")
```


# 2018 category levels


Percent of Vehicles Available

```{r}
jsonData <- read_json(vehicles_url, simplifyVector = TRUE)

vehicles <- jsonData$features$attributes %>%
select(OBJECTID, CSA2010, novhcl18, Shape__Area, Shape__Length)

vehicle_table_for_joining <- vehicles %>%
  mutate(vehicle_boolean = ifelse(novhcl18 > 30, 1, 0)) %>%
  select(CSA2010, vehicle_boolean, percnt_no_vehicle = novhcl18)
head(vehicle_table_for_joining)

```
```{r}
jsonData <- read_json(income_url, simplifyVector = TRUE)

med_income <- jsonData$features$properties %>% 
  select(OBJECTID, CSA2010, mhhi18, Shape__Area, Shape__Length)

income_table_for_joining <- med_income %>%
  mutate(income_boolean = ifelse(mhhi18 < 46435, 1, 0)) %>%
  select(CSA2010, income_boolean, median_income = mhhi18)
```
```{r}
master_neighbourhood_data_2018 <- income_table_for_joining %>%
  left_join(vehicle_table_for_joining, by = 'CSA2010') %>%
  left_join(grocery_store_join, by = 'CSA2010') %>%
  left_join(hfai_for_joining, by = 'CSA2010') %>%
  mutate(
    number_criteria_met = income_boolean + vehicle_boolean + has_store + hfai_boolean,
    priority_level = case_when(
      number_criteria_met == 4 ~ 'High',
      number_criteria_met > 2 ~ 'Medium',
      number_criteria_met < 3 ~ 'Low'
    )
  )
master_neighbourhood_data_2018
```

```{r}
write_csv(master_neighbourhood_data_2018, "master_neighbourhood_data_2018.csv")
```


# Loading demographic data from API

## Age

### Percent of Population 5-17 years old

```{r}
library(jsonlite)
library(tidyverse)

pop_children_url <- "https://services1.arcgis.com/mVFRs7NF4iFitgbY/arcgis/rest/services/Age18_/FeatureServer/0/query?where=1%3D1&outFields=*&outSR=4326&f=json"

jsonData <- read_json(pop_children_url, simplifyVector = TRUE)

pop_children <- jsonData$features$attributes %>% 
  select(OBJECTID, CSA2010, age18_22, Shape__Area, Shape__Length)

head(pop_children)
```

### Percent of Population 18-24 years old

```{r}
library(jsonlite)
library(tidyverse)

pop_youngadult_url <- "https://services1.arcgis.com/mVFRs7NF4iFitgbY/arcgis/rest/services/Age24_/FeatureServer/0/query?where=1%3D1&outFields=*&outSR=4326&f=json"

jsonData <- read_json(pop_youngadult_url, simplifyVector = TRUE)

pop_youngadult <- jsonData$features$attributes %>% 
  select(OBJECTID, CSA2010, age24_22, Shape__Area, Shape__Length)

head(pop_youngadult)
```

### Percent of Population 24-64 years old

```{r}
library(jsonlite)
library(tidyverse)

pop_adult_url <- "https://services1.arcgis.com/mVFRs7NF4iFitgbY/arcgis/rest/services/Age64_/FeatureServer/0/query?where=1%3D1&outFields=*&outSR=4326&f=json"

jsonData <- read_json(pop_adult_url, simplifyVector = TRUE)

pop_adult <- jsonData$features$attributes %>% 
  select(OBJECTID, CSA2010, age64_22, Shape__Area, Shape__Length)

head(pop_adult)
```

### Percent of Population 65 and older

```{r}
library(jsonlite)
library(tidyverse)

pop_elderly_url <- "https://services1.arcgis.com/mVFRs7NF4iFitgbY/arcgis/rest/services/Age65_/FeatureServer/0/query?where=1%3D1&outFields=*&outSR=4326&f=json"

jsonData <- read_json(pop_elderly_url, simplifyVector = TRUE)

pop_elderly <- jsonData$features$attributes %>% 
  select(OBJECTID, CSA2010, age65_22, Shape__Area, Shape__Length)

head(pop_elderly)
```

## Race

### Racial Diversity Index

The percent chance that two people picked at random within an area will be of a different race/ethnicity. This number does not reflect which race/ethnicity is predominant within an area. The higher the value, the more racially and ethnically diverse an area.

```{r}
library(jsonlite)
library(tidyverse)

rdi_url <- "https://services1.arcgis.com/mVFRs7NF4iFitgbY/arcgis/rest/services/Racdiv/FeatureServer/0/query?where=1%3D1&outFields=*&outSR=4326&f=json"

jsonData <- read_json(rdi_url, simplifyVector = TRUE)

rdi <- jsonData$features$attributes %>% 
  select(OBJECTID, CSA2010, racdiv22, Shape__Area, Shape__Length)

head(rdi)
```

### Percent of Residents All Other Races

The percentage of persons, out of the total number of persons living in an area, who self-identify as either American Indian, Alaskan Native, Native Hawaiian or Other Pacific Islander, or some other race (non-Hispanic).

```{r}
library(jsonlite)
library(tidyverse)

pop_other_url <- "https://services1.arcgis.com/mVFRs7NF4iFitgbY/arcgis/rest/services/Ppac/FeatureServer/0/query?where=1%3D1&outFields=*&outSR=4326&f=json"

jsonData <- read_json(pop_other_url, simplifyVector = TRUE)

pop_other <- jsonData$features$attributes %>% 
  select(OBJECTID, CSA2010, ppac22, Shape__Area, Shape__Length)

head(pop_other)
```

### Percent of Residents Black/African-American

```{r}
library(jsonlite)
library(tidyverse)

pop_black_url <- "https://services1.arcgis.com/mVFRs7NF4iFitgbY/arcgis/rest/services/Paa/FeatureServer/0/query?where=1%3D1&outFields=*&outSR=4326&f=json"

jsonData <- read_json(pop_black_url, simplifyVector = TRUE)

pop_black <- jsonData$features$attributes %>% 
  select(OBJECTID, CSA2010, paa22, Shape__Area, Shape__Length)

head(pop_black)
```

### Percent of Residents White (Caucasian)

```{r}
library(jsonlite)
library(tidyverse)

pop_white_url <- "https://services1.arcgis.com/mVFRs7NF4iFitgbY/arcgis/rest/services/Pwhite/FeatureServer/0/query?where=1%3D1&outFields=*&outSR=4326&f=json"

jsonData <- read_json(pop_white_url, simplifyVector = TRUE)

pop_white <- jsonData$features$attributes %>% 
  select(OBJECTID, CSA2010, pwhite22, Shape__Area, Shape__Length)

head(pop_white)
```

### Percent of Residents Hispanic

```{r}
library(jsonlite)
library(tidyverse)

pop_hispanic_url <- "https://services1.arcgis.com/mVFRs7NF4iFitgbY/arcgis/rest/services/Phisp/FeatureServer/0/query?where=1%3D1&outFields=*&outSR=4326&f=json"

jsonData <- read_json(pop_hispanic_url, simplifyVector = TRUE)

pop_hispanic <- jsonData$features$attributes %>% 
  select(OBJECTID, CSA2010, phisp22, Shape__Area, Shape__Length)

head(pop_hispanic)
```

### Percent of Resident Asian (Non-Hispanic)

```{r}
library(jsonlite)
library(tidyverse)

pop_asian_url <- "https://services1.arcgis.com/mVFRs7NF4iFitgbY/arcgis/rest/services/Pasi/FeatureServer/0/query?where=1%3D1&outFields=*&outSR=4326&f=json"

jsonData <- read_json(pop_asian_url, simplifyVector = TRUE)

pop_asian <- jsonData$features$attributes %>% 
  select(OBJECTID, CSA2010, pasi22, Shape__Area, Shape__Length)

head(pop_asian)
```

### Percent Population (25 Years and over) With Less Than a High School Diploma or GED

```{r}
library(jsonlite)
library(tidyverse)

pop_no_hs_degree <- "https://services1.arcgis.com/mVFRs7NF4iFitgbY/arcgis/rest/services/Lesshs/FeatureServer/0/query?where=1%3D1&outFields=*&outSR=4326&f=json"

jsonData <- read_json(pop_no_hs_degree, simplifyVector = TRUE)

pop_no_hs_degree <- jsonData$features$attributes %>% 
  select(OBJECTID, CSA2010, lesshs21, Shape__Area, Shape__Length)

head(pop_no_hs_degree)
```

### Percent Population (25 Years and over) With High School Diploma and Some College or Associates Degree

```{r}
library(jsonlite)
library(tidyverse)

pop_hs_degree <- "https://services1.arcgis.com/mVFRs7NF4iFitgbY/arcgis/rest/services/Hsdipl/FeatureServer/0/query?where=1%3D1&outFields=*&outSR=4326&f=json"

jsonData <- read_json(pop_hs_degree, simplifyVector = TRUE)

pop_hs_degree <- jsonData$features$attributes %>% 
  select(OBJECTID, CSA2010, hsdipl21, Shape__Area, Shape__Length)

head(pop_hs_degree)
```

### Joining demographic data with master neighbourhood data

```{r}
demo_data <- cbind(pop_children$CSA2010, pop_children$age18_22, pop_youngadult$age24_22, pop_adult$age64_22, pop_elderly$age65_22, rdi$racdiv22, pop_other$ppac22, pop_black$paa22, pop_white$pwhite22, pop_hispanic$phisp22, pop_asian$pasi22, pop_no_hs_degree$lesshs21, pop_hs_degree$hsdipl21)
demo_data <- as.data.frame(demo_data)
colnames(demo_data) <- c("CSA2010", "age18_22", "age24_22", "age64_22", "age65_22", "racdiv22", "ppac22", "paa22", "pwhite22", "phisp22", "pasi22", "lesshs21", "hsdipl21")
# combine demo_data and master_data
demo_master_data <- left_join(master_neighbourhood_data, demo_data, by = "CSA2010")
write_csv(demo_master_data, "demo_master_data.csv")
```


# Demographics for areas with over 30% of households without a vehicle

```{r}
library(dplyr)

# filter demographic datasets
## percent of population 5-17 years old
vehicles_30_children <- pop_children %>% 
  filter(CSA2010 %in% vehicles_30$CSA2010) %>%
  select(CSA2010, age18_22)

## percent of population 18-24 years old
vehicles_30_youngadult <- pop_youngadult %>% 
  filter(CSA2010 %in% vehicles_30$CSA2010) %>%
  select(CSA2010, age24_22)

## percent of population 24-64 years old
vehicles_30_adult <- pop_adult %>% 
  filter(CSA2010 %in% vehicles_30$CSA2010) %>%
  select(CSA2010, age64_22)

## percent of population 65 and older
vehicles_30_elderly <- pop_elderly %>% 
  filter(CSA2010 %in% vehicles_30$CSA2010) %>%
  select(CSA2010, age65_22)

## racial diversity index
vehicles_30_rdi <- rdi %>% 
  filter(CSA2010 %in% vehicles_30$CSA2010) %>%
  select(CSA2010, racdiv22)

## percent of residents all other races
vehicles_30_other <- pop_other %>% 
  filter(CSA2010 %in% vehicles_30$CSA2010) %>%
  select(CSA2010, ppac22)

## percent of residents black/african-american
vehicles_30_black <- pop_black %>% 
  filter(CSA2010 %in% vehicles_30$CSA2010) %>%
  select(CSA2010, paa22)

## percent of residents white
vehicles_30_white <- pop_white %>% 
  filter(CSA2010 %in% vehicles_30$CSA2010) %>%
  select(CSA2010, pwhite22)

## percent of residents hispanic
vehicles_30_hispanic <- pop_hispanic %>% 
  filter(CSA2010 %in% vehicles_30$CSA2010) %>%
  select(CSA2010, phisp22)

## percent of residents asian
vehicles_30_asian <- pop_asian %>% 
  filter(CSA2010 %in% vehicles_30$CSA2010) %>%
  select(CSA2010, pasi22)

# merge datasets
vehicles_30_demo <- vehicles_30_children %>% 
  full_join(vehicles_30_youngadult, by = "CSA2010") %>% 
  full_join(vehicles_30_adult, by = "CSA2010") %>% 
  full_join(vehicles_30_elderly, by = "CSA2010") %>% 
  full_join(vehicles_30_rdi, by = "CSA2010") %>% 
  full_join(vehicles_30_other, by = "CSA2010") %>% 
  full_join(vehicles_30_black, by = "CSA2010") %>% 
  full_join(vehicles_30_white, by = "CSA2010") %>% 
  full_join(vehicles_30_hispanic, by = "CSA2010") %>% 
  full_join(vehicles_30_asian, by = "CSA2010")

head(vehicles_30_demo)
summary(vehicles_30_demo)
```

# Building dashboards with flexdashboard

## Install and load packages

```{r}
#install.packages(c("flexdashboard", "DT", "shiny"))
library(flexdashboard)
library(DT)
library(shiny)
```


## Address Lookup Code  

```{r}
library(tidygeocoder)
library(sf)
library(geosphere)

nearest_store <- function(address, city, state, zip) {
   if (missing(address) | missing(city) | missing(state) | missing(zip)){
    stop('Please enter a valid address including street address, city, state, and zip code.')
  }
  address_to_geocode <- data.frame(address = 
                      paste(address, ", ", city, ", ", state, " ", zip, sep = ""))
 
  coordinates <- suppressMessages(tidygeocoder::geocode(address_to_geocode, address, method = "osm"))
  address.coords <- data.frame(
    lats = coordinates$lat,
    lons = coordinates$long)
  
  grocery.store.coords <-
   data.frame(
    name = grocery_stores$storename,
    address_groc = grocery_stores$address,
    lats_ = grocery_stores$latitude,
    lons_ = grocery_stores$longitude)
  
  new <- grocery.store.coords %>% rowwise() %>%
  mutate(new =  map(list(c(lons_, lats_)), ~ address.coords %>% rowwise() %>%
                      mutate(dist = distGeo((c(lons, lats)), .x)) %>%
                      ungroup %>%
                      filter(dist == min(dist))))  %>%
  ungroup() %>% unnest_wider(new) %>% 
  filter(dist == min(dist)) 
  
  distance <- round(new$dist*0.000621371,2)
  
    paste0("Your nearest grocery store is ", new$name, " at ", new$address_groc, ".", " It is ", distance, " miles from your inputted address, ", paste(address, ", ", city, ", ", state, " ", zip, sep = ""), ".")
}

# testing function
nearest_store("615 N Wolfe St")
```

```{r}

 map(.x = asplit(grocery_stores[,2:3], 1), .f = ~ distGeo(c(coordinates$long, coordinates$lat), .x))

  address_to_geocode <- data.frame(address = 
                      "3250 E Fayette St", "Baltimore", "MD", "21224")
  coordinates <- tidygeocoder::geocode(address_to_geocode, address, method = "osm")
  map(.x = asplit(grocery_stores[,3:2],1), .f = ~ distGeo(p1 = matrix(c(coordinates$long, coordinates$lat),byrow=T,nrow=1), p2 = .x))
  
list(c(grocery_stores$longitude, grocery_stores$latitude))

asplit(grocery_stores[,3:2],1)

point_sf <- st_as_sf(coordinates, coords = c('long', 'lat'))
  point_sf
  grocery_store_list <- st_as_sf(grocery_stores, coords = c('longitude', "latitude"))
  st_nearest_feature(point_sf, grocery_store_list)
  st_nearest_feature
st_nearest_feature(event.coords.sf, locale.centroids.sf)

map(c(grocery_stores$longitude, grocery_stores$latitude), distGeo(p1 = c(coordinates$long, coordinates$lat), p2 = .x))


coordinates %>% 
  distGeo(c(long, lat), c())

event.coords %>% rowwise() %>%
  mutate(new =  map(list(c(lons, lats)), ~ locale.centroids %>% rowwise() %>%
                      mutate(dist = distGeo((c(lons, lats)), .x)) %>%
                      ungroup %>%
                      filter(dist == min(dist)) %>%
                      select(locale.id, dist)))  %>%
  ungroup() %>% unnest_wider(new)
```





