---
title: "Food Accessibility in Baltimore by Neighborhood"
author: "Amia Graye, Rosie Quint, and Julia Martin"
output: 
  flexdashboard::flex_dashboard: 
    social: menu
    source_code: embed
    vertical_layout: fill
    orientation: columns
    style: css
runtime: shiny
---

```{r global, include=FALSE}
# load data in 'global' chunk so it can be shared by all users of the dashboard
master_url <- "https://raw.githubusercontent.com/jhu-statprogramming-fall-2024/project4-apicons/refs/heads/main/master_neighbourhood_data.csv"
master_data <- read.csv(master_url)
```

```{r setup, include=FALSE}
library(flexdashboard)
library(DT)
library(shiny)
library(ggplot2)
library(dplyr)
library(tidyverse)
library(jsonlite)
library(mapbaltimore)
library(leaflet)
library(tidygeocoder)
library(sf)
library(geosphere)
library(bslib)
```

# Welcome

### Welcome 

```{r}
h1(strong("Welcome!"), style = "font-size:30px;")
```


This is a project created by three Biostatistics ScM students at Johns Hopkins University. We are exploring food accessibility in Baltimore by neighborhood. We will be looking at the distribution of grocery stores in Baltimore, the demographic characteristics of neighborhoods, and the accessibility of grocery stores in Baltimore. This is a resource for residents of Baltimore to find their nearest grocery store and learn more about their neighborhood, or for policymakers who are seeking a broad overview of food accessibility in Baltimore.

# Categorization of Baltimore Neighborhoods

Map of Baltimore neighborhoods with priority levels

```{r}
#Getting neighbourhood boundary geometries
csa_name_geometry <- mapbaltimore::csas %>%
  rename(CSA2010 = name)

# Adding geometries to master neighbourhood data 
master_data_with_geometry <- master_data %>% 
  left_join(csa_name_geometry, by = "CSA2010") 

#Turn into SF object
master_data_with_geometry <- st_as_sf(master_data_with_geometry)

# Transform CRS to 4326 (needed to interpret the multipolygon values)
master_data_with_geometry <- st_transform(master_data_with_geometry, crs = 4326)

# define factor for colours 
master_data_with_geometry$priority_level <- factor(
  master_data_with_geometry$priority_level,
  levels = c("High", "Medium", "Low")
)

#Set colour levels (for low medium and high categories) 
pal <- colorFactor(
  palette = c("red", "yellow", "green"), 
  domain = master_data_with_geometry$priority_level
)

#Plot map with colour coded neighbourhoods by priority level
leaflet(master_data_with_geometry) %>%
  addTiles() %>%  # Add base map tiles
  addPolygons(
    color = "blue",        # Border color
    weight = 1,            # Border thickness
    fillColor = ~pal(priority_level), # Fill color
    fillOpacity = 0.7,     # Transparency of the fill
    popup = ~paste0(
      "<strong>Neighborhood:</strong> ", CSA2010, "<br>",
      "<strong>Median Income:</strong> $", format(median_income, big.mark = ","), "<br>",
      "<strong>Percent Without Vehicle:</strong> ", round(percnt_no_vehicle, 1), "%<br>",
      "<strong>Has Grocery Store:</strong> ", ifelse(has_store == 1, "Yes", "No")
    ),
    label = ~paste0(
      CSA2010, "<br>",
      "Median Income: $", format(median_income, big.mark = ","), "<br>",
      "No Vehicle: ", round(percnt_no_vehicle, 1), "%<br>",
      "Grocery Store: ", ifelse(has_store == 1, "Yes", "No")
    ),
    labelOptions = labelOptions(
      style = list("font-weight" = "bold"), # Bold text
      textsize = "12px",                   # Font size
      direction = "auto"                   # Automatically adjust label placement
    )
  ) %>%
  addLegend(
    position = "bottomright",
    colors = c("red", "yellow", "green"),         # Explicit color assignment
    labels = c("High", "Medium", "Low"),          # Explicit labels in desired order
    opacity = 0.7,
    title = "Neighbourhood Healthy Food Priority Level"
  )
```

# Find Your Nearest Grocery Store

### Address Lookup

```{r}
#define address lookup function

grocery_store_url <- "https://services1.arcgis.com/UWYHeuuJISiGmgXx/arcgis/rest/services/Grocery_Store/FeatureServer/0/query?where=1%3D1&outFields=*&outSR=4326&f=json"

jsonData <- read_json(grocery_store_url, simplifyVector = TRUE)

grocery_stores <- jsonData$features$attributes

nearest_store <- function(address, city, state, zip) {
   if (address == "" | city == ""){
    stop('Please enter a valid address including at minimum street address and city.')
  }
  address_to_geocode <- data.frame(address = 
                      paste(address, ", ", city, ", ", state, " ", zip, sep = ""))
 
  coordinates <- suppressMessages(tidygeocoder::geocode(address_to_geocode, address, method = "osm"))
  address.coords <- data.frame(
    lats = coordinates$lat,
    lons = coordinates$long)
  
  grocery.store.coords <-
   data.frame(
    name = grocery_stores$storename,
    address_groc = grocery_stores$address,
    lats_ = grocery_stores$latitude,
    lons_ = grocery_stores$longitude)
  
  new <- grocery.store.coords %>% rowwise() %>%
  mutate(new =  map(list(c(lons_, lats_)), ~ address.coords %>% rowwise() %>%
                      mutate(dist = distGeo((c(lons, lats)), .x)) %>%
                      ungroup %>%
                      filter(dist == min(dist))))  %>%
  ungroup() %>% unnest_wider(new) %>% 
  filter(dist == min(dist)) 
  
  distance <- round(new$dist*0.000621371,2)
  
  first <- paste0("Your nearest grocery store is ", new$name, " at ", new$address_groc, ".")
  
  second <- paste0("It is ", distance, " miles from your inputted address, ", paste(address, ", ", city, ", ", state, " ", zip, sep = ""), ".")
  
   return(c(first, second))
}

coords <- function(address, city, state, zip) {
  address_to_geocode <- data.frame(address = 
                      paste(address, ", ", city, ", ", state, " ", zip, sep = ""))
 
  coordinates <- suppressMessages(tidygeocoder::geocode(address_to_geocode, address, method = "osm"))
  
    address.coords <- data.frame(
    lats = coordinates$lat,
    lons = coordinates$long)
  
  grocery.store.coords <-
   data.frame(
    name = grocery_stores$storename,
    address_groc = grocery_stores$address,
    lats_ = grocery_stores$latitude,
    lons_ = grocery_stores$longitude)
  
  new <- grocery.store.coords %>% rowwise() %>%
  mutate(new =  map(list(c(lons_, lats_)), ~ address.coords %>% rowwise() %>%
                      mutate(dist = distGeo((c(lons, lats)), .x)) %>%
                      ungroup %>%
                      filter(dist == min(dist))))  %>%
  ungroup() %>% unnest_wider(new) %>% 
  filter(dist == min(dist)) 
  
  leaflet() %>%
  addTiles() %>%
  addMarkers(lng=new$lons_, lat=new$lats_, popup=new$name)
}
```

```{r}
h1(strong("Find your nearest grocery store!"), style = "font-size:30px;")

ui <- fluidPage(
  sidebarPanel(
  textInput('address', "Street Address", value = "", width = NULL, placeholder = NULL),
textInput('city', "City", value = "", width = NULL, placeholder = NULL),
textInput('state', "State", value = "", width = NULL, placeholder = NULL),
textInput('zip', "Zip Code", value = "", width = NULL, placeholder = NULL),
actionButton("go", "Go"),
  textOutput("add"), width = 5
),
mainPanel(
  leafletOutput("map"), width = 7
)
)

server <- function(input, output) {
  
  store <- eventReactive(input$go, {nearest_store(input$address, input$city, input$state, input$zip)})
  c <- eventReactive(input$go, {coords(input$address, input$city, input$state, input$zip)})

  output$add <- renderText(store())
  output$map <- renderLeaflet(c())
}

shinyApp(ui, server)
```

### Map of Baltimore grocery stores
```{r}
# Create the leaflet map
leaflet(master_data_with_geometry) %>%
  addTiles() %>% # Add base map tiles
  addPolygons(
    color = "black",                 
    weight = 1,                       
    fillColor = "transparent",       
    fillOpacity = 0,                
    popup = ~paste0("<strong>Neighborhood:</strong> ", CSA2010), 
    label = ~CSA2010,                 
    labelOptions = labelOptions(
      style = list("font-weight" = "bold"), 
      textsize = "12px",
      direction = "auto"
    )
  ) %>%
  addMarkers(
    data = grocery_stores,     
    lng = ~longitude,          
    lat = ~latitude,                
    popup = ~paste0(
      "<strong>Store Name:</strong> ", storename, "<br>",
      "<strong>Address:</strong> ", address)
  )
```
# Demographic Characteristics of Baltimore Neighborhoods
